---
title: "Predicting the brexit vote"
author: ""
date: "25/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data

- In this exercise, we will work on a classification task of Brexit referendum vote
- The data is originally from British Election Study Online Panel
  - codebook: https://www.britishelectionstudy.com/wp-content/uploads/2020/05/Bes_wave19Documentation_V2.pdf
- The outcome is `LeaveVote` (1: Leave, 0: Remain)

## Libraries

- We will use the following packages

```{r}
library(tidyverse)
library(caret)
library(glmnet)
```

## Load data

We sub-sample the data. Full data takes too much time to estimate for the class... (Feel free to run full sample after the class)

```{r}
set.seed(20200813)
data_brexit <- read_csv("data/data_bes.csv.gz") %>%
  sample_n(3000) # sampling data so
```


## Data preparation

- We will carry out:
  - make `LeaveVote` factor variable
  - test train split
  - preprocess


```{r}
data_brexit <- data_brexit %>%
    mutate(LeaveVote = factor(LeaveVote))
```

### Train-test split

```{r}
train_idx <- createDataPartition(data_brexit$LeaveVote, p = .7, list = F) 

data_train <- data_brexit %>% slice(train_idx)
data_test <- data_brexit %>% slice(-train_idx)
```

### Preprocess

```{r}
prep <- preProcess(data_train %>% select(-LeaveVote), method = c("center", "scale"))
prep

data_train_preped <- predict(prep, data_train)
data_test_preped <- predict(prep, data_test)

```

## Model formulas

There are four logistic regression models  in the manuscript (Table 2).

1. Sociodemographics
2. Identity
3. Anti-elite
4. Attitudes

The following line of codes will generate the each model. 

```{r}
fm_socdem <- formula("LeaveVote ~ gender + age + edlevel + hhincome + econPersonalRetro1")
fm_identity <- formula("LeaveVote ~ gender + age + edlevel + hhincome + 
                        EuropeanIdentity + EnglishIdentity + BritishIdentity")
fm_antielite <- formula("LeaveVote ~ gender + age + edlevel + hhincome + 
              PolMistrust + GovDisapproval + PopulismScale + 
              ConVote + LabVote + LibVote + SNPPCVote + UKIP")
fm_attitudes <- formula("LeaveVote ~ gender + age + edlevel + hhincome + euUKNotRich + 
              euNotPreventWar + FreeTradeBad + euParlOverRide1 + euUndermineIdentity1 + 
              lessEUmigrants + effectsEUTrade1 + effectsEUImmigrationLower")
fm_all <- formula("LeaveVote ~ .")


```

You can use these formulas in a way like:

```{r eval = F}
# for model
glm(fm_socdem, data = data_train_preped, family = "binomial")
# for data extraction
model.matrix(fm_socdem, data = data_train_preped) %>% head()

```

## Logistic regression

Run a few models, and evaluate them. Which one has the better predictive performance?

```{r}
mod <- glm(fm_socdem, data = data_train_preped, family = "binomial")
test_pred <- (predict(mod, newdata = data_test_preped) > .5) %>% 
  as.numeric() %>% factor()
confusionMatrix(test_pred, data_test_preped$LeaveVote, 
                positive = "1", mode = "prec_recall")

```
```{r}
print_test_confusion_matrix <- function(fm) {
  mod <- glm(fm, data = data_train_preped, family = "binomial")
  test_pred <- (predict(mod, newdata = data_test_preped) > .5) %>% 
    as.numeric() %>% factor()
  confusionMatrix(test_pred, data_test_preped$LeaveVote, 
                  positive = "1", mode = "prec_recall")
}

print_test_confusion_matrix(fm_socdem)
print_test_confusion_matrix(fm_identity)
    
print_test_confusion_matrix(fm_antielite)
print_test_confusion_matrix(fm_attitudes)
print_test_confusion_matrix(fm_all)
```
## Linear SVM

- Train a linear SVM model, check the predictive performance. How does it compare to the logistic regression?

```{r}
model_svm <- train(fm_identity, data = data_train_preped, method = "svmLinear")
print(model_svm)

test_pred <- predict(model_svm, newdata = data_test_preped) 
confusionMatrix(test_pred, data_test_preped$LeaveVote, 
                positive = "1", mode = "prec_recall")


train_model <- function(fm, method = "svmLinear") {
  model <- train(fm, data = data_train_preped, method = method)
  return(model)
}

formulae <- list(fm_socdem, fm_identity, fm_antielite, fm_attitudes)

models <- formulae %>% map(train_model)
models %>% map(print)
```
```{r}
models %>% 
  map(~predict(., newdata = data_test_preped)) %>%
  map(~confusionMatrix(., data_test_preped$LeaveVote, positive = "1", mode = "prec_recall"))
```


## Polynomial SVM and Radial SVM

- Train non-linear SVM. How is the performance? Any improvement?

```{r cache=T}

model_svm_poly <- train(fm_identity, data = data_train_preped, method = "svmPoly")

model_svm_poly
pred_test_svm_poly <- predict(model_svm_poly, newdata = data_test_preped)
confusionMatrix(pred_test_svm_poly, data_test_preped$LeaveVote, positive = "1", mode = "prec_recall")

models_svmradial <- formulae %>% map(~train_model(., method = "svmRadial"))
models_svmradial %>% map(print)

models_svmradial %>% 
  map(~predict(., newdata = data_train_preped)) %>%
  map(~confusionMatrix(., data_train_preped$LeaveVote, positive = "1", mode = "prec_recall"))

models_svmradial %>% 
  map(~predict(., newdata = data_train_preped)) %>%
  map(~confusionMatrix(., data_train_preped$LeaveVote)$byClass %>% data.frame()) %>%
  bind_cols()


```


## (Optional) Logistic regression with LASSO

- `glmnet` can run a Logistic model with L1 penalty (LASSO). 
- Try a "full" model combining all inputs.
  - Which inupts survived?

```{r}
library(glmnet)
train_y <- data_train_preped$LeaveVote

#train_x <- data_train_preped %>% select(-LeaveVote) %>% as.matrix()
train_x <- model.matrix(fm_all, data = data_train_preped)[, -1]


model_lasso_cv <- cv.glmnet(train_x, train_y,
                             data = data_train_preped,
                             alpha = 1,
                             family = "binomial")
plot(model_lasso_cv)

```

```{r}
plot(model_lasso_cv$glmnet.fit, xvar = "lambda", label = T)
colnames(train_x)

coef(model_lasso_cv)
```
```{r}
test_x <- model.matrix(fm_all, data = data_test_preped)[, -1]
pred_test_lasso <- (predict(model_lasso_cv, newx = test_x) > .5) %>%
  as.numeric() %>% factor()

confusionMatrix(pred_test_lasso, data_test_preped$LeaveVote, 
                positive = "1", mode = "prec_recall")
```